Looking for a Signal in the Noise: Revisiting Obesity and the Microbiome

### 2016-07-20 submission/Sze_Obesity_mBio_2016.Rmd
Move Supplemental Text to Methods

***Literature Review and Study Inclusion.***
To perform a robust meta-analysis and limit inclusion bias, we followed the Preferred Reporting Items for Systematic Reviews and Meta-Analyses (PRISMA) guidelines to identify the studies that we analyzed [@moher_preferred_2010]. A detailed description of our selection process and the exact search terms are provided in the Supplemental Text and in Figure 1. Briefly, we searched PubMed for original research studies that involved studying obesity and the human microbiome. The initial search yielded 187 studies. We identified 10 additional studies that were not designed to explicitly test for an association between the microbiome and obesity. We then manually curated the 197 studies to select those studies that included BMI and 16S rRNA gene sequence data. This yielded 11 eligible studies. An additional study was removed from our analysis because no individuals in the study had a BMI over 30. Among the final 10 studies, 3 were identified from our PubMed search [@turnbaugh_obesity-associated_2006; @zupancic_analysis_2012; @escobar_gut_2014], 5 were originally identified from the 10 studies that did not explicitly investigate obesity but included BMI data [@zeevi_personalized_2015; @ross_16s_2015; @goodrich_human_2014; @wu_linking_2011; @human_microbiome_project_consortium_structure_2012], and two datasets were used [@schubert_microbiome_2014; @baxter_microbiota-based_2016] because these publications did not specifically look for any metabolic or obesity conditions but had control populations and enabled us to help mitigate against publication biases associated with the bacterial microbiome and obesity. The `r number_string[n]` studies are summarized in Table 1. For comparison, two of these studies were included in the Finucane re-analysis study [@turnbaugh_obesity-associated_2006; @human_microbiome_project_consortium_structure_2012] and four of these studies were included in the Walters re-analysis study [@turnbaugh_obesity-associated_2006; @human_microbiome_project_consortium_structure_2012; @zupancic_analysis_2012; @wu_linking_2011]. The 16S rRNA gene sequence data from each study was re-analyzed using a similar approach based on previously described methods for reducing the number of chimeric sequences and sequencing errors for 454 and Illumina MiSeq data [@Kozich_2013; @Schloss2011]. The sequences were clustered into operational taxonomic units (OTUs) using the average neighbor approach [@Westcott2015] and into taxonomic groupings based on their classification using a naive Bayesian classifier [@Wang2007].

@@

>>>>

***Literature Review and Study Inclusion.***
To perform a robust meta-analysis and limit inclusion bias, we followed the Preferred Reporting Items for Systematic Reviews and Meta-Analyses (PRISMA) guidelines to identify the studies that we analyzed [@moher_preferred_2010]. A detailed description of our selection process and the exact search terms are provided in the Methods section and in Figure 1. Briefly, we searched PubMed for original research studies that involved studying obesity and the human microbiome. The initial search yielded 187 studies. We identified 10 additional studies that were not designed to explicitly test for an association between the microbiome and obesity. We then manually curated the 197 studies to select those studies that included BMI and 16S rRNA gene sequence data. This yielded 11 eligible studies. An additional study was removed from our analysis because no individuals in the study had a BMI over 30. Among the final 10 studies, 3 were identified from our PubMed search [@turnbaugh_obesity-associated_2006; @zupancic_analysis_2012; @escobar_gut_2014], 5 were originally identified from the 10 studies that did not explicitly investigate obesity but included BMI data [@zeevi_personalized_2015; @ross_16s_2015; @goodrich_human_2014; @wu_linking_2011; @human_microbiome_project_consortium_structure_2012], and two datasets were used [@schubert_microbiome_2014; @baxter_microbiota-based_2016] because these publications did not specifically look for any metabolic or obesity conditions but had control populations and enabled us to help mitigate against publication biases associated with the bacterial microbiome and obesity. The `r number_string[n]` studies are summarized in Table 1. For comparison, two of these studies were included in the Finucane re-analysis study [@turnbaugh_obesity-associated_2006; @human_microbiome_project_consortium_structure_2012] and four of these studies were included in the Walters re-analysis study [@turnbaugh_obesity-associated_2006; @human_microbiome_project_consortium_structure_2012; @zupancic_analysis_2012; @wu_linking_2011]. The 16S rRNA gene sequence data from each study was re-analyzed using a similar approach based on previously described methods for reducing the number of chimeric sequences and sequencing errors for 454 and Illumina MiSeq data [@Kozich_2013; @Schloss2011]. The sequences were clustered into operational taxonomic units (OTUs) using the average neighbor approach [@Westcott2015] and into taxonomic groupings based on their classification using a naive Bayesian classifier [@Wang2007].

@@

***In-Depth Overview of Search Strategy.***
The initial search strategy included looking for all papers that initially fit under the below NCBI PubMed advanced search criteria: "(((((((((Bacterial Microbiome) AND (Obesity or bmi or body mass index or BMI or obesity) AND "last 10 years"[PDat] AND Humans[Mesh])) NOT review[ptyp]) AND "last 10 years"[PDat] AND Humans[Mesh])) AND "last 10 years"[PDat] AND Humans[Mesh])) AND "last 10 years"[PDat] AND Humans[Mesh])". The terms included in this criteria were that the manuscript had to have "Bacterial Microbiome" and "Obesity, BMI, bmi, obesity" in their manuscript criteria, it was not published more than 10 years ago, they were not review articles, and it contained research on humans only. This search yielded a total of 187 manuscripts. From the Finucane and Walters re-analysis studies along with knowledge of other published papers that included BMI information in their sequence metadata we obtained 10 additional studies. This brought our total number of records to 197. We browsed abstracts of these studies and included studies that mentioned stool or feces examination, did not involve children, was not a clinical trial for probiotics or other diet related treatments, did not have participants with inflammatory bowel disease, were in English, and were based on the analysis of sequence data from the 454 or MiSeq platforms.  This ultimately excluded all but a total of 11 studies. Because we decided *a priori* to use the standard definition for BMI group classification one study was excluded because it did not have any individuals who were obese.
